### 2018-05-15 General WLD combinatorics.tex
Fixed a problem and proved the next part of the annoying proof about the properties of alg 4.1

The final thing to prove for this item is that there are no irreducible factors $f$ of degree 3 or more.  Now suppose for a contradiction that $f$ has degree $\geq 3$. Note that by removing the propagators which do not contribute to $f$ and changing $i$ to be the first vertex which contributes to $f$ we obtain a different admissible diagram for which $f=\Delta_{I_i}$ and $i\in I_i$.  Showing that this different admissible diagram gives a contradiction is sufficient, and so we may assume that $W$ has $\Delta_{I_i}$ irreducible of degree 3 or more and $i\in I_i$.

@@

\textbf{Case I}: Suppose $p$ has one end on the edge $(i-1, i)$.  Thus $p$ is supported on $(i-1, i, m, m+1)$ for some $m >_i i$.  Then $I_{i+1}(p) = m$ by Lemma~\ref{lem susama}.  Call the propagators which appear before $p$ is closed the propagators inside $p$. \note{It might be nice to have general language for this inside/outside idea.  What do you think.}  The construction of $I_{i+1}$ begins as for $I_i$ after the first step; they can only differ once $p$ contributes to $I_{i+1}$.  Thus if a propagator contributes $m$ in $I_i$ then it cannot be a propagator inside $p$ (nor $p$ itself).  Let $S$ be the set of propagators inside $p$ along with $p$ itself.  Two things can now happen; if neither $m$ nor $m+1$ appear in $I_i$ then the only vertex in the support of $p$ in $I_i$ is $i$, and so the row of $p$ in the matrix of $\Delta_{I_i}$ has only one nonzero entry; hence $\Delta_{I_i}$ has a linear factor which is a contradiction.  The other think which can happen is that at least one of $m$ and $m+1$ is in $I_i$.  However, all the propagators in $S$ are mapped by the bijection of $I_{i}$ to vertices strictly before $m$ so the matrix giving $\Delta_{I_i}$ has the form
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the matrix for the propagators in $S$ and the vertices in $I_i(S)$ which is square of size $|S|\times |S|$. No other propagators can be supported on these vertices since all other propagators are outside of $p$ and $p$ is the first propagator supported at $i$.  This explains the zero block.  Therefore $\Delta_{I_i} = \det A \det C$ and both factors are nontrivial since $m$ appears in $I_i$.  This contradicts the irreducibility of $\Delta_{I_i}$.

\textbf{Case II}: Suppose $p$ has one end on the edge $(i, i+1)$.  If no other propagators are supported on $i$ then the column for vertex $i$ has only one nonzero entry in it and so $\Delta_{I_i}$ has a linear factor which is a contradiction.  Thus we can take $q$ to be the propagator such that $I_i(q)=i+1$ and we know that $q$ has one end on the edge $(i, i+1)$ and is adjacent to $p$ on that edge in the counterclockwise direction.  Write $(i, i+1, j, j+1)$ for the support of $q$.  The situation for $q$ is very similar to case I, in particular $I_{i+1}(q) = j$ by Lemma~\ref{lem susama} and so if $j\in I_i$ then the propagator which contributes it is outside of $q$.  Thus similarly to case I, if we let $S$ be the set of propagators inside $q$ along with $p$ and $q$ themselves then all propagators in $S$ are mapped by $I_i$ to vertices strictly before $j$ and no other propagators cover vertices strictly before $j$.  Thus the matrix giving $\Delta_{I_i}$ has the form
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the matrix for the propagators in $S$ and the vertices in $I_i(S)$. Again two things can now happen.  If some vertex $j$ or larger in $>_i$ appears in $I_i$ then $B$ and $C$ have at least one column and so the block form of the matrix gives a nontrivial factorization of $\Delta_{I_i}$ which is a contradiction.  If no vertex $j$ or larger is in $I_i$ then the matrix giving $\Delta_{I_i}$ is just $A$.  Looking in more detail into $A$, note that of supports of $p$ and $q$, only $i$ and $i+1$ are in $I_i$ and hence
\[
A = \begin{bmatrix} D & 0 \\ E & F\end{bmatrix}
\]
where $D$ is the $2\times 2$ matrix indexed by the propagators $p$ and $q$ and the vertices $i$ and $i+1$.  By assumption $\Delta_{I_i}$ had degree 3 or more and so this matrix decomposition gives that $\Delta_{I_i}$ factors into nontrivial factors contradicting our assumptions.

@@

    \item
    \item
    \item
  \end{enumerate}
\end{proof}

>>>>

The final thing to prove for this item is that there are no irreducible factors $f$ of degree 3 or more.  Now suppose for a contradiction that $f$ has degree $\geq 3$. Note that by removing the propagators which come before those contributing to $f$ and changing $i$ to be the first vertex which contributes to $f$ we obtain a different admissible diagram for which $f$ still divides $\Delta_{I_i}$ but also $i\in I_i$ and $i$ contributes to $f$.  Showing that this different admissible diagram gives a contradiction is sufficient, and so we may assume that $i\in I_i$ and $i$ contributes to $f$.  Finally, we can suppose that $W$ is minimal in number of propagators with the above occuring.

@@

\textbf{Case I}: Suppose $p$ has one end on the edge $(i-1, i)$.  Thus $p$ is supported on $(i-1, i, m, m+1)$ for some $m >_i i$.  Then $I_{i+1}(p) = m$ by Lemma~\ref{lem susama}.  Call the propagators which appear before $p$ is closed the propagators inside $p$. \note{It might be nice to have general language for this inside/outside idea.  What do you think.}  The construction of $I_{i+1}$ begins as for $I_i$ after the first step; they can only differ once $p$ contributes to $I_{i+1}$.  Thus if a propagator contributes $m$ in $I_i$ then it cannot be a propagator inside $p$ (nor $p$ itself).  Let $S$ be the set of propagators inside $p$ along with $p$ itself.  Two things can now happen; if neither $m$ nor $m+1$ appear in $I_i$ then the only vertex in the support of $p$ in $I_i$ is $i$, and so the row of $p$ in the matrix of $\Delta_{I_i}$ has only one nonzero entry; hence $\Delta_{I_i}$ has a linear factor contributed by $p$ and $i$ which is a contradiction.  The other thin which can happen is that at least one of $m$ and $m+1$ is in $I_i$.  However, all the propagators in $S$ are mapped by the bijection of $I_{i}$ to vertices strictly before $m$ so the matrix giving $\Delta_{I_i}$ has the form
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the matrix for the propagators in $S$ and the vertices in $I_i(S)$ which is square of size $|S|\times |S|$. No other propagators can be supported on these vertices since all other propagators are outside of $p$ and $p$ is the first propagator supported at $i$.  This explains the zero block.  Therefore $\Delta_{I_i} = \det A \det C$ and both factors are nontrivial since $m$ appears in $I_i$.  If we remove the propagator outside of $p$ we get a smaller diagram for which $\Delta_{I_i} = \det A$, this contradicts the minimality of our choices unless $\det A$ is quadratic which in turn contradicts that $i$ and $p$ contribute to an irreducible factor $f$ of degree at least 3.

\textbf{Case II}: Suppose $p$ has one end on the edge $(i, i+1)$.  If no other propagators are supported on $i$ then the column for vertex $i$ has only one nonzero entry in it and so $\Delta_{I_i}$ has a linear factor which is a contradiction.  Thus we can take $q$ to be the propagator such that $I_i(q)=i+1$ and we know that $q$ has one end on the edge $(i, i+1)$ and is adjacent to $p$ on that edge in the counterclockwise direction.  Write $(i, i+1, j, j+1)$ for the support of $q$.  The situation for $q$ is very similar to case I, in particular $I_{i+1}(q) = j$ by Lemma~\ref{lem susama} and so if $j\in I_i$ then the propagator which contributes it is outside of $q$.  Thus similarly to case I, if we let $S$ be the set of propagators inside $q$ along with $p$ and $q$ themselves then all propagators in $S$ are mapped by $I_i$ to vertices strictly before $j$ and no other propagators cover vertices strictly before $j$.  Thus the matrix giving $\Delta_{I_i}$ has the form
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the matrix for the propagators in $S$ and the vertices in $I_i(S)$. Again two things can now happen.  If some vertex $j$ or larger in $>_i$ appears in $I_i$ then $B$ and $C$ have at least one column and so the block form of the matrix gives a nontrivial factorization of $\Delta_{I_i}$.  Hence, as in case I either this contradicts minimality by removing those propagators not contributing to $A$, or $\det A$ is quadratic which contradicts that $p$ and $i$ contribute to $f$, an irreducible factor of degree at least 3.  On the other hand, if no vertex $j$ or larger is in $I_i$ then the matrix giving $\Delta_{I_i}$ is just $A$.  Looking in more detail into $A$, note that of supports of $p$ and $q$, only $i$ and $i+1$ are in $I_i$ and hence
\[
A = \begin{bmatrix} D & 0 \\ E & F\end{bmatrix}
\]
where $D$ is the $2\times 2$ matrix indexed by the propagators $p$ and $q$ and the vertices $i$ and $i+1$.  Thus $p$ and $i$ contribute to a quadratic factor of $\Delta_{I_i}$ contradicting our assumptions.

@@

    \item Suppose propagators $p$ and $q$ are supported on a common edge $(a,b)$ and $I_i(p)=a$, $I_i(q)=b$.  For any fixed bijection $\sigma$ of the propagators other than $p$ and $q$ to $I_i -\{a,b\}$ for which each propagator is supported on its image under the bijection, we can extend $\sigma$ to a bijection of all propagators with $I_i$ in two ways, either $p\rightarrow a$ and $q\rightarrow b$ or $p\rightarrow b$ and $q\rightarrow a$.  The sum of the contributions of all these bijections to $\Delta_{I_i}$ is the product of the minor of $C(W)$ coming from $p, q$ and $a,b$ with the minor coming from the other propagators and $I_i - \{a,b\}$.  Since there is no cancellation of terms in the expansion of $\Delta_{I_i}$, if any other terms appear then they will cause a factor which is not in the form described in the previous part.  Therefore no such terms exist and $x_1x_4-x_2x_3$ is a factor of $\Delta_{I_i}$.

Now let $f$ be a quadratic factor of $\Delta_{I_i}$.  By the previous part we know that $f$ is a $2\times 2$ minor coming from two propagators, call them $p$ and $q$, and two vertices, call them $a <_i b$.  It remains to show that $a$ and $b$ are adjacent.  From this we can conclude that $p$ and $q$ each have one end on $(a,b)$, as any other way for both $p$ and $q$ to be supported on two consecutive vertices would contradict noncrossing or the density requirement of admissibility.

As in the proof of the previous part, make a new admissible diagram by removing the propagators which come before $f$ and set $i=a$.  The cases in the proof of the previous part show how $\Delta_{I_i}$ factors; in particular the vertices covering the second end of $p$ either do not appear in $I_i$ or contribute to a different factor of $\Delta_{I_i}$ than $p$ and $a$ do.  By assumption $b$ contributes to the same factor as $a$.  Therefore $(a,b)$ is an edge.
    \item
    \item
  \end{enumerate}
\end{proof}
